---
title: "nyc_002"
output: html_notebook
---
For asynchronous lecture on Sept 11, 2023.


Task1: Loading necessary packages:
```{r}

library(nycflights13)
library(tidyverse)
library(janitor)
library(here)
rm()

df1 = nycflights13::flights

```

Task2: Catching up to prior output (flight 83 delay plot)
```{r}

df6 = df1 |> filter(flight == 83, dest == "MIA")

ggplot(df6, aes(x=dep_delay, y=arr_delay)) + geom_point()

# Questions:
# Do delays change by time of day?
# Are delays more frequent at certain times of year?
# Do flights "hurry up" in response to dep_delay?

```

Task3: Flights from LGA to MIA, observing delay response
```{r}

df2 = df1 |> filter(dest == "MIA") |> count(origin, sort = TRUE)
view(df2)

df7 = df1 |> filter(dest == "MIA", origin == "LGA") |> mutate(flt_delta = arr_delay - dep_delay)

ggplot(df7, aes(x=dep_delay, y=flt_delta)) + geom_point(alpha=0.1)

```
It appears that flt_delta is more often negative than positive, and the more severe the dep_delay, the more negative the delta, meaning they hurry up to make up time.


Task4: Using horizontal line indicator of average delta
```{r}

ggplot(df7, aes(x=dep_delay, y=flt_delta)) + geom_point(alpha=0.1) + geom_hline(aes(yintercept = mean(flt_delta, na.rm = TRUE)))

```

Task5: Seeing how dep_delay depends on time of year
```{r}

df7 |> 
  ggplot(aes(x=time_hour,y=dep_delay)) + geom_point(alpha=0.1) + stat_smooth() + ylim(-25,120)

```

Task6: Seeing how dep_delay depends on time of day
```{r}
# attempt1
df7 |> ggplot(aes(x=hour+minute/60,y=dep_delay)) + geom_point(alpha=0.1) + stat_smooth() + ylim(-25,120)

# attempt2 --> more detailed
df7 |> mutate(day_of_week=weekdays(time_hour)) |> ggplot(aes(x=hour+minute/60,y=dep_delay, color=day_of_week)) + geom_point(alpha=0.1) + stat_smooth() + ylim(-25,120)

# attempt3 --> most detailed
df7 |> mutate(day_of_week=weekdays(time_hour)) |> ggplot(aes(x=hour+minute/60,y=dep_delay, color=day_of_week)) + geom_point(alpha=0.1) + stat_smooth() + ylim(-20,40) + facet_wrap(~day_of_week)

```

Task7: Starting new BTS data study
```{r}

# Loading the BTS data looks different for me here because I unzipped the file so I am loading the csv directly

BTSdata = here("rawdata","9.11_BTS_data","T_T100D_SEGMENT_US_CARRIER_ONLY.csv")

df2022 = read_csv(BTSdata) |> clean_names()

```
Task 8: Extracting useful observations, plotting flights by month from LGA
```{r}
df9 = df2022 |> filter(passengers != 0, origin == "LGA")

# bar chart counting the number of rows by month
ggplot(df9, aes(month)) + geom_bar()

# bar chart counting number of departures by month
ggplot(df9, aes(month)) + geom_bar(aes(weight=departures_performed))

# bar chart counting number of passengers by month
ggplot(df9, aes(month)) + geom_bar(aes(weight=passengers))
```
Observation: Fewer passengers early 2022 due to COVID19 still being prominent.

Task 9: Further analysis...
```{r}

# Coloring previous figure by carrier
ggplot(df9, aes(month, fill = carrier_name)) + geom_bar(aes(weight=departures_performed))


# Plotting total passengers through LGA
df10 = df2022 |> filter(passengers>0, origin == "LGA" | dest == "LGA")

ggplot(df10, aes(month)) + geom_bar(aes(weight=passengers))

# Decluttering variables
df11 = df10 |> select(month,passengers,carrier_name,dest,origin)

df12 = df10 |> select(1:5,month, contains("id"))

```

Task 10: How full was the plane?
```{r}

df13 = df10 |> mutate(percent_loading = passengers/seats*100)

ggplot(df13, aes(percent_loading)) + geom_histogram() + facet_wrap(~carrier_name, scales = "free_y")

```


```{r}
